---
title: 'Single HiC QC report'
date: "`r date()`"
author: 'HiCool'
format:
    html:
        css: "%CSS%"
        highlight-style: github
        code-block-bg: true
        code-block-border-left: "#3e3e3e"
        toc: true
        toc-depth: 2
        self-contained: true
        embed-resources: true
execute:
    echo: false
---

```{r return = FALSE, output = FALSE}
## -- Load required libraries
require(HiCExperiment)
require(HiCool)

## -- Parse data 
file_cool <- "%COOL%"
file_pairs <- "%PAIRS%"
file_log <- "%LOG%"
list_args <- getHiCoolArgs(file_log)
list_stats <- getHicStats(file_log)

## -- Recreate cf
x <- CoolFile(
    path = file_cool, 
    pairs = file_pairs, 
    metadata = list(
        log = file_log, 
        stats = list_stats, 
        args = list_args
    )
)

## -- Generate strings for text 
string_hicool <- paste0("HiCool v. ", packageVersion('HiCool'))
string_r1 <- list_args[['r1']]
string_r2 <- list_args[['r2']]
string_genome <- list_args[['genome']]
string_folder <- list_args[['wd']]
log_lines <- readLines(file_log)
string_date_start <- gsub(".*: ", "", grep('## date: ', log_lines, value = TRUE))
string_date_end <- gsub(
    ",", " ", 
    gsub(" ::.*", "", log_lines[tail(which(grepl('^[0-9][0-9][0-9][0-9]-[0-9][0-9]-[0-9][0-9],', log_lines)), 1)])
)
.dhms <- function(t) {
    paste(
        t %/% (60*60*24), ' days, ',
        paste(
            formatC(t %/% (60*60) %% 24, width = 2, format = "d", flag = "0"), 'h ',
            formatC(t %/% 60 %% 60, width = 2, format = "d", flag = "0"), 'min ',
            formatC(t %% 60, width = 2, format = "d", flag = "0"), 's ',
            sep = ""
        )
    )
}
string_runtime <- .dhms(as.numeric(difftime(string_date_end, string_date_start)))
```

## Processing

\ **Run time**  
\ Started @ `r string_date_start`  
\ Finished @ `r string_date_end`  
\ Duration: `r string_runtime`

\ **Processing directory**  
\ ```r string_folder```  

\ **Reads location**  
\ R1: ```r string_r1```  
\ R2: ```r string_r2```  

\ **Genome reference**  
\ ```r string_genome``` 

**Command**  

```{r}
if (length(list_args)) {
    df_args <- lapply(
        seq_len(length(list_args)-1), 
        function(K) {
            data.frame(
                arg = names(list_args)[K], 
                value = ifelse(
                    is.character(list_args[[K]]), 
                    paste0('"', list_args[[K]], '"'), 
                    as.character(list_args[[K]])
                )
            )
        }
    )
    df_args <- do.call(rbind, df_args)
    message(
        "x <- HiCool::HiCool(\n", 
        paste0(apply(df_args, 1, function(x) {paste0("\t", x[1], " = ", x[2])}), collapse = ',\n'), 
        "\n)"
    )
} else {
    message("HiCool arguments not found. ")
}
```

\ **Output**
```{r}
x
```

<hr>

## Mapping results

| | |
| ------------------------------------------------------------------- | ---------------------------------------------------------------------------------- |
| **Number of paired-end fragments**                                  | `r formatC(list_stats$nFragments, format = "f", big.mark = ",", digits = 0)`     |
| **Total number of pairs after mapping**                             | `r paste0(formatC(list_stats$nPairs, format = "f", big.mark = ",", digits = 0))`         |
| **Number of valid pairs**                                           | `r paste0(formatC(list_stats$nFiltered, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nFiltered / list_stats$nPairs, 4) * 100, '% of all pairs)')`         |
| **Number of unique valid pairs**                                    | `r paste0(formatC(list_stats$nUnique, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nUnique / list_stats$nPairs, 4) * 100, '% of all pairs)')`         |
| Number of pairs from uncut fragments (+-)                           | `r paste0(formatC(list_stats$nDangling, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nDangling / list_stats$nPairs, 4) * 100, '% of all pairs)')`         |
| Number of pairs from self-religated fragments (-+)                  | `r paste0(formatC(list_stats$nSelf, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nSelf / list_stats$nPairs, 4) * 100, '% of all pairs)')`         |
| Number of dumped pairs (-- or ++)                                   | `r paste0(formatC(list_stats$nDumped, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nDumped / list_stats$nPairs, 4) * 100, '% of all pairs)')`         |

\ Minimum # of restriction sites to retain \`+-\` pairs: `r list_stats[['threshold_uncut']]`  
\ Minimum # of restriction sites to retain \`-+\` pairs: `r list_stats[['threshold_self']]`  

```{r "Camembert plot"}
df <- data.frame(
    type = c(
        "Valid pairs (unique)", 
        "Valid pairs (duplicates)", 
        'Dangling fragments ("uncut")', 
        'Self-religated fragments ("loops")', 
        'Invalid pairs ("dumped")'  
    ), 
    score = c(
        list_stats$nUnique, 
        list_stats$nDups, 
        list_stats$nDangling, 
        list_stats$nSelf, 
        list_stats$nDumped
    ), 
    text = c(
        paste0("Valid pairs (unique):\n", formatC(list_stats$nUnique, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nUnique / list_stats$nPairs, 4) * 100, '% of all pairs)'),
        paste0("Valid pairs (duplicates):\n", formatC(list_stats$nDups, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nDups / list_stats$nPairs, 4) * 100, '% of all pairs)'),
        paste0('Dangling fragments ("uncut"):\n', formatC(list_stats$nDangling, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nDangling / list_stats$nPairs, 4) * 100, '% of all pairs)'),
        paste0('Self-religated fragments ("loops"):\n', formatC(list_stats$nSelf, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nSelf / list_stats$nPairs, 4) * 100, '% of all pairs)'),
        paste0('Invalid pairs ("dumped"):\n', formatC(list_stats$nDumped, format = "f", big.mark = ",", digits = 0), ' (', round(list_stats$nDumped / list_stats$nPairs, 4) * 100, '% of all pairs)')
    )
)
colors <- c('#D35E60', '#f3b6b7', '#5ec9d3', '#5ed371', '#d3b25e')
fig <- plotly::plot_ly(
    df, labels = ~type, values = ~score, type = 'pie',
    textposition = 'inside',
    textinfo = 'percent',
    insidetextfont = list(color = '#FFFFFF'),
    hoverinfo = 'text',
    text = ~text,
    marker = list(
        colors = colors,
        line = list(color = '#FFFFFF', width = 3), 
        pull = 0.1
    ),
    showlegend = TRUE
)
fig <- fig |> plotly::layout(
    title = paste0(
        'Distribution of pairs in \n', 
        gsub('.pairs', '', basename(pairsFile(x))), ' sample'
    ), 
    margin = list(t = 150)
)
fig
```

## Hi-C quality control

```{r output = FALSE}
if (require(HiContacts)) {
    hcf <- import(file_cool, pairsFile = file_pairs)
    ps <- HiContacts::distanceLaw(hcf, by_chr = TRUE) |> suppressMessages()
    fig <- plotly::plot_ly(
        ps, 
        x = ~binned_distance, y = ~norm_p, 
        type = 'scatter', mode = 'lines', 
        color = ~chr,
        showlegend = TRUE
    )
    fig <- fig |> plotly::layout(
        title = paste0(
            "P(s) of ", gsub('.pairs', '', basename(pairsFile(x))), " valid pairs"
        ), 
        xaxis = list(
            title = 'Distance between interacting loci (s)', 
            type = "log"
        ),
        yaxis = list(
            title = 'Normalized P(s)', 
            type = "log"
        )
    )
}
```

```{r}
if (require(HiContacts)) {
    fig
} else {
    message("Install HiContacts package to compute distance law and % of trans contacts")
    message("`BiocManager::install('HiContacts')`")
}
```

cis/trans per chr.

## Session information

> System

```{r}
message(
    "OS:\t\t\t\t", sessioninfo::os_name(), '\n',
    "system:\t\t\t", version$system, '\n',
    "UI:\t\t\t\t", .Platform$GUI, '\n',
    "R:\t\t\t\t", paste(
        version$major, version$minor, sep = '.'
    ), "(", R.home(), ")", '\n',
    "Python:\t\t\t", basilisk::listPythonVersion(HiCool:::env_HiCool), "(", file.path(
        basilisk::obtainEnvironmentPath(HiCool:::env_HiCool), 'bin', 'python'
    ), ")", '\n',
    "Bioconductor:\t", BiocManager::version(), '\n',
    "basilisk:\t\t", packageVersion('basilisk'), '\n',
    "HiCExperiment:\t", packageVersion('HiCExperiment'), '\n',
    "HiCool:\t\t\t", packageVersion('HiCool'), '\n'
)
```

> System packages

```{r}
message(
    "samtools: \t", gsub('.*==', '', grep('samtools', HiCool:::.HiCool_dependencies, value = TRUE)), "\n",
    "bowtie2: \t", gsub('.*==', '', grep('bowtie2', HiCool:::.HiCool_dependencies, value = TRUE)), "\n"
)
```

> R packages 

```{r}
sessioninfo::package_info()
```

> Python environment

```{r}
proc <- basilisk::basiliskStart(HiCool:::env_HiCool)
sessioninfo::python_info()
basilisk::basiliskStop(proc)
```

> Python packages

```{r}
basilisk::listPackages(HiCool:::env_HiCool)
```
